%%\documentclass{article}
%%\documentclass[12pt,a4paper]{article}
\documentclass[12pt,reqno,a4paper]{amsart}

% --- preamble --- 


% 1. Encoding and Fonts (standard)
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

% 2. Math and standard AMS packages
\usepackage{amsmath, amsfonts, amssymb, amsthm}

% more packages

\usepackage{listings}
\lstset{ %
	breaklines=true, %
	backgroundcolor=\color{white}, %
	commentstyle=\color{green}, %
	keywordstyle=\color{blue}\textbf,  %
stringstyle=\color{orange}, %
identifierstyle=\color{red}\itshape,  %
}
\usepackage{color}
\usepackage{imakeidx}
% generate an index file:
\makeindex


% 3. Metadata and Linking (Load these LAST)
\usepackage[hidelinks]{hyperref} % 'hidelinks' removes the ugly red/green boxes
\usepackage{hyperxmp}

\hypersetup{
%%    pdftitle={Example LaTeX article
    pdfauthor={Author Name},
    pdfcopyright={Licensed under CC BY 4.0},
    pdflicenseurl={https://creativecommons.org/licenses/by/4.0/}
}

% 4. Optional: Simple Footer License
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt} % Remove the horizontal line at the top
\fancyfoot[C]{\thepage}
\fancyfoot[L]{\footnotesize CC BY 4.0}

% --- article metadata ( titlepage info) ---

\title[short title]{example \LaTeX\ article}
\author{Author Name}
%  Standard amsart: email and address print at the very end
% \email{\href{mailto:mame@mail.domain}{mailto:name@mail.domain}}
% Optional: Use \thanks if you want the email on page 1
% %\thanks{Correspondence: email@example.com}
% \thanks{Correspondence: \href{mailto:mame@mail.domain}{mailto:name@mail.domain}}
\thanks{Correspondence: \href{mailto:mame@mail.domain}{mailto:name@mail.domain}}
\address{The address}
\date{\today}


% --- document body ---
\begin{document}

\begin{abstract}

This is a simple  example article document to  help getting started with  {\LaTeX}.

\end{abstract}
%% note: for amsart, abstact must be before maketitle
\maketitle

\cleardoublepage
\phantomsection % This creates a precise anchor for the PDF link
\addcontentsline{toc}{section}{Contents}
\tableofcontents

\addcontentsline{}{section}{Introduction}
\section*{Introduction}

This is an example document written in {\LaTeX}, Leslie Lamport's\index{Lamport, Leslie} typesetting
sysyem which is based on Donald E. Knuth's\index{Knuth, Donald E.} {\TeX}.


The main philosphy behind {\LaTeX}---in contrast to most word processors\index{word processor} --- is to allow and encourage the author to
focus on the content of the text rather than on the layout\index{layout}and formatting.


writing documents in {\LaTeX} is a bit like programming\index{programming}: One writes a ``source code''\index{source code}  file with the content in plain text
interspersed with some {\LaTeX} commands\index{commands}, also called macros.\index{macros} .\footnote{{\LaTeX} is in fact a collection of {\TeX} macros.}\label{fn:macros}
and then ``compiles'' or processes it with the processorss \verb|latex| or \verb|pdflatex| to produce
putput files in formats like \textbf{PDF}\index{pdf} or {\TeX}' own device independent format \textbf{DVI}\index{dvi}.

It is not the purpose of this document to be a comprehensive guide but rather to serve as an example demonstating some of the capabilities of  {\LaTeX}.


some more introductory words.


And this is already the end of the introduction.




\section{Beginnings}
\label{sec:beginnings}

This first section explains and demonstrates some concepts of {\LaTeX}.

\subsection{paragraphs}

Paragraphs are automatically created from blocks of text separated by one or more (tradittionally two or more)
blank lines.


Some more text goes here in a new paragraph.
If starting a new paragraph by adding blank lines does not work (which can be the case for some versions of {\LaTeX}),  a new paragraph can also be started explicitly with the \verb|\paragraph{}| command.
Alternatively, a paragraph can be ended with the \verb|\par| command.


Paragraphs are automatically separated by a vertical space of height \verb|\parskip| and the first line of each paragraph is by default indented by \verb|\parindent|



\subsection{structuring elements}

Section headers and other structure elements are automatically formatted and numbered
if one uses the corresponding commands like \verb|\section|, \verb|\subsection|,  \verb|\chapter|, \verb|\part|, and
more.


Of these, chaper and part are typically only used for longer documents like books while sections and subsections are the typical structure of articles.


In addition, \verb|\paragraph| can also create numbered paragraphs with titles but this is rarely used and might be confusing for the overall structure of the document.



\subsection{math formulae}

 As {\LaTeX} is based on Donald E. Knuth's\index{Knuth} {\TeX},
Writing math\index{math} formulae\index{formulae} is rather easy. in the \verb+equation+ environment.

Here is an example:


\begin{equation}
\label{eq:example}
\sum_{i=1}^{n}i = \frac{n(n+1)}{2}
\end{equation}


Note that the  previous equation is numered and can thus be referenced (see below). Alternatively, formulae can be written inline (without numbering), by surrounding them
with \verb|$| characters. for example $a^2 + b^2 = c^2$.

\section{Continuation}

This section covers a few more interesting aspects of {\LaTeX}.


\subsection{Creating a table of contents}


One big strength of {\LaTeX} is the ability to automatically create a table of contents from the
structure elements. The conmand \verb+\tableofcontents+ generates the contents using  information gathered from the structure elements likesection, subsection, chapter, etc.


In order to get the page references right, \verb+latex+ (or \verb+pdflatex+, respectively) has to be run at least twice on the input file.


If needed, contents entries can also  be added manually with the \verb+\addcontentsline+ command.
This can be useful in order to have unnumbered structure elements like \verb+\section*+ show up in the contents.

\subsection{Labels And References}


Another very convenient feature is the ability to reference\index{reference}  sections and other numerable elements like formulae
and images in the text. An item can be referenced with the \verb+\ref+ command if it has been given a label\index{label} with the \verb+\label+ command.  Example: Section \textbf{Beginnings} has number \ref{sec:beginnings}.
Equation \ref{eq:example} is the numbered example equation.

\subsection{Footnotes}


It is furthermore very easy to create professional looking numbered  footnotes\index{footnote}\footnote{example footnote} in \LaTeX\ with the \verb+\footnote+ command.\footnote{another footnote}
 





\section{Indexes}

Another extremely useful feature\index{feature} of  {\LaTeX} ---especially for longer documents like books\index{books}--- is the ability to automate\index{automate} the creation of indexes\index{indexes} of keywords.

Using the package \verb+makeidx+ or \verb+imakeidx+ and placing the  command  \verb+\makeindex+ in the document''s preamble, on can use the command \verb+\index{keyword}+ in the text to add the keyword\index{keyword}  \texttt{keyword} to the index.

\verb+\index+ can be used anywhere in the text but will typically be used near an  appearance of the keyword.

When creating an index,  The comman \verb+makeindex+ has to be run between runs of \verb+latex+ an \verb+pdflatex+, respectively.

I. E. with an input file named \verb+doc.tex+, a typical compile\index{compile} sequence could look like this:


%%\begin{verbatim}

\begin{lstlisting}
 # pdflatex doc
 # pdflatex doc
 # makeindex doc
 # pdflatex doc
\end{lstlisting}
%%\end {verbatim}


\section{Code listings}

source code for various languages  can be conveniently listed an nicely highlighted with the \verb+lstlisting+ environment provided by the \verb+listings+ package.


 As an example, here is the \verb+Makefile+ for this example document:

\lstset{language=[gnu] make}
\begin{lstlisting}

SHELL=/bin/sh

.SUFFIXES: .tex .pdf

BASE=doc
SOURCE=$(BASE).tex
TARGET=$(BASE).pdf

FILES_WC=$(wildcard $(BASE).*)
FILES_LS=$(shell ls $(BASE).*)
FILES=$(FILES_LS)

GENERATED_S=$(subst $(SOURCE), ,$(FILES))
GENERATED_FO=$(filter-out $(SOURCE),$(FILES))
GENERATED=$(GENERATED_FO)

.PHONY: info

info:
	@echo       source: $(SOURCE)
	@echo       target: $(TARGET)
	@echo all files wc: $(FILES_WC)
	@echo all files ls: $(FILES_LS)
	@echo    all files: $(FILES)
	@echo  gen-files-s:  $(GENERATED_S)
	@echo gen-files-fo: $(GENERATED_FO)
	@echo    gen-files: $(GENERATED)



.tex.pdf:
	@pdflatex $<
	@pdflatex $<
	@makeindex $<
	@pdflatex $<
	@pdflatex $<

.PHONY: $(BASE)
	
$BASE): pdf
.PHONY: pdf
pdf: clean $(TARGET)

$(TARGET): $(SOURCE)

show:  $(BASE)
	@$(VIEWER) $(TARGET)

.PHONY: clean
clean:
	rm -f $(GENERATED)

\end{lstlisting}






\section{Conclusion}

With this, our little example document already comes to an end.



%%\section*{Index}
% load and print  the index:
\printindex
\addcontentsline{}{section}{Index}


\end{document}
